---
title: "Checking Lab Range"
subtitle: "MS Registry DQ Work"
author: "Ziyu Huang"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: readable
    toc: true
    toc_float: true
    number_sections: true
  word_document:
    toc: true
    toc_depth: 2
---



```{r set-options, echo=FALSE, cache=FALSE}
# Clear environment
rm(list=ls())

# Chunk options
knitr::opts_chunk$set(comment=NA, echo=FALSE)

# Some colors

```


```{r packages, message=FALSE}
# 
# 
# library(skimr)
# 

# library(broom)
#library(Epi)
# 

# library(lme4)

## Load Love-boost then the packages
# source(here("R", "Love-boost.R"))
# library(here)
# library(survival)
# library(survminer)
# library(magrittr)
#library(data.table)
#library(tableone)
# library(naniar)
# library(janitor)
library(readxl)
library(openxlsx)
library(ggrepel)
library(haven)
library(tidyverse)

theme_set(theme_bw())

```

```{r paths and parameters}
myinput = "~/../../Corrona LLC/Biostat Data Files - MS/monthly/2024/2024-05-05/final/ms_dmlabimg_2024-05-05.dta"
myoutput = NULL
myspecs = "~/../../Corrona LLC/Biostat Data Files - MS/DQ checks/specs/ms_labtest_range.xlsx"
myresults = "~/../../Corrona LLC/Biostat Data Files - MS/DQ checks/reports/ms_dmlabimg nc00 labtest range checks.xlsx"
```


```{r data import}
df.raw <- read_dta(myinput)
df.specs <- read_xlsx(myspecs)
df.out <- read_xlsx(myresults)
```


# Summary

This is to examine the MS lab and imaging data and help make a decision where some variables might have ambiguous "standard units".

# Variables


## Lymphocytes


Here's a summary of the results recorded for Lymphocytes.

```{r}
# create temp data frame for exploration
df <- df.raw %>% 
  filter(event_type_dec == "Lymphocytes") %>% 
  select(id, result)

# get summary statistics
mosaic::favstats(df$result) %>% 
  knitr::kable()


# How many are over 100
summary_df <- df %>%
  mutate(category = ifelse(result <= 100, "Under or Equal to 100", "Over 100")) %>%
  count(category) %>%
  mutate(percentage = n / sum(n) * 100) 


```

Here, only `r summary_df[1,2]` observations out of `r nrow(df)` is over 100, which accounts for `r round(summary_df[1,3],2)`%.

```{r}
summary_df %>% knitr::kable(digits = 2)

```

```{r}
summary_df <- df %>%
  mutate(category = case_when(
    result <= 10 ~ "0 to 10",
    result > 10 & result <= 65 ~ "10 to 65",
    result > 65 & result <= 100 ~ "65 to 100",
    result > 100 ~ "Over 100"
  )) %>%
  count(category) %>%
  mutate(percentage = n / sum(n) * 100)

summary_df %>% knitr::kable(digits = 2)
```

Setting a lower bound of range check to be 10% means we will see `r summary_df[1,2]` observations in the report, which accounts for `r round(summary_df[1,3],2)`% of the observations.

## Summary of other variables

```{r}
within_bounds <- function(value, bounds) {
  bounds <- str_remove_all(bounds, "\\[|\\]") %>% str_split(",") %>% unlist()  %>% str_trim()%>% as.numeric()
  lower <- bounds[1]
  upper <- bounds[2]
  return(value >= lower & value <= upper)
}

# Split df.specs into two parts
df.specs_with_dec <- df.specs %>% filter(!is.na(event_type_dec))
df.specs_no_dec <- df.specs %>% filter(is.na(event_type_dec))

# Join df.raw with df.specs_with_dec
report_with_dec <- df.raw %>%
  inner_join(df.specs_with_dec, by = c("eventtypedesc", "event_type_dec")) %>%
  rowwise() %>%
  mutate(within_bounds = within_bounds(result.x, result.y)) %>%
  ungroup() %>%
  mutate(failure = if_else(within_bounds, 0, 1, missing = 1))

# Join df.raw with df.specs_no_dec using only eventtypedesc
report_no_dec <- df.raw %>%
  inner_join(df.specs_no_dec, by = "eventtypedesc") %>%
  rowwise() %>%
  mutate(within_bounds = within_bounds(result.x, result.y)) %>%
  ungroup() %>%
  mutate(failure = if_else(within_bounds, 0, 1, missing = 1))

# Combine the two reports
combined_report <- bind_rows(report_with_dec, report_no_dec)

# Summarise the combined report
final_report <- combined_report %>%
  group_by(eventtypedesc, event_type_dec) %>%
  summarise(
    failures = sum(failure, na.rm = TRUE), 
    total_observations = n(),
    .groups = "drop"
  ) %>% 
  mutate(
    temp = failures / total_observations,
    percent_failure = glue::glue("{round((failures / total_observations) * 100, 1)}%")
  ) %>% 
  arrange(desc(temp)) %>% 
  select(-temp)
knitr::kable(final_report)
```

## Jitter plots


```{r}
# Create a list to store the plots
plot_list <- list()

# Get the unique event types
event_types <- unique(df.out$event_type_dec)

# Iterate through each event type and create a jitter plot with horizontal lines and labels next to the y-axis
for (event in event_types) {
  # Filter the data for the current event type
  event_data <- df.out %>% filter(event_type_dec == event)
  
  # Create the jitter plot
  plot <- ggplot(event_data, aes(x = event_type_dec, y = result)) +
    geom_jitter(width = 0.2, height = 0, alpha = 0.5) +
    geom_hline(aes(yintercept = lowerBound), color = "blue", linetype = "dashed") +
    geom_hline(aes(yintercept = upperBound), color = "red", linetype = "dashed") +
    geom_text(aes(x = 0.6, y = lowerBound, label = paste("Lower Bound:", lowerBound)),
              vjust = 1, color = "blue", size = 4) +
    geom_text(aes(x = 1.4, y = upperBound, label = paste("Upper Bound:", upperBound)),
               vjust = -1, color = "red", size = 4) +
    labs(title = paste("Jitter Plot of Result for Test Type:", event),
         x = "Test Type",
         y = "Result") +
    theme_minimal() +
    theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  
  # Add the plot to the list
  plot_list[[event]] <- plot
}

# Display all plots
for (event in event_types) {
  print(plot_list[[event]])
}



```

```{r}

# Filter out rows where failure is 1 (out of bounds)
out_of_bounds <- combined_report %>%
  filter(failure == 1) %>%
  arrange(eventtypedesc, event_type_dec, result.x)

# Generate a list of data frames, one for each event type
out_of_bounds_list <- out_of_bounds %>%
  group_by(eventtypedesc, event_type_dec) %>%
  group_split()

# Optionally, name each data frame in the list for easy reference
names(out_of_bounds_list) <- out_of_bounds %>%
  group_by(eventtypedesc, event_type_dec) %>%
  group_keys() %>%
  mutate(name = paste(eventtypedesc, event_type_dec, sep = "_")) %>%
  pull(name)

```


```{r}
# Load the necessary package


# Create a new workbook
wb <- createWorkbook()

# Function to truncate sheet names to a maximum of 31 characters
truncate_sheet_name <- function(name) {
  if (nchar(name) > 31) {
    return(substr(name, 1, 31))
  } else {
    return(name)
  }
}

# Add each data frame in out_of_bounds_list to a new sheet
for (name in names(out_of_bounds_list)) {
  # Clean and truncate the sheet name to ensure it's valid
  sheet_name <- str_replace_all(name, "[^[:alnum:]_]", "_")
  sheet_name <- truncate_sheet_name(sheet_name)
  
  # Select only the required columns
  selected_columns <- out_of_bounds_list[[name]] %>%
    select(eventtypedesc, event_type_dec, result.x, result.y)
  
  # Add a new sheet to the workbook
  addWorksheet(wb, sheet_name)
  
  # Write the data frame with selected columns to the sheet
  writeData(wb, sheet_name, selected_columns)
}



# Save the workbook to a file
saveWorkbook(wb, "../output/out_of_bounds_report.xlsx", overwrite = TRUE)


```

# Session Information
```{r session information, include=T}
sessionInfo()
```

