---
title: "Checking Lab Range"
subtitle: "MS Registry DQ Work"
author: "Ziyu Huang"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: readable
    toc: true
    toc_float: true
    number_sections: true
  word_document:
    toc: true
    toc_depth: 2
---



```{r set-options, echo=FALSE, cache=FALSE}
# Clear environment
rm(list=ls())

# Chunk options
knitr::opts_chunk$set(comment=NA, echo=FALSE)

# Some colors

```


```{r packages, message=FALSE}
# 
# 
# library(skimr)
# 

# library(broom)
#library(Epi)
# 

# library(lme4)

## Load Love-boost then the packages
# source(here("R", "Love-boost.R"))
# library(here)
# library(survival)
# library(survminer)
# library(magrittr)
#library(data.table)
#library(tableone)
# library(naniar)
# library(janitor)
# library(readxl)
library(haven)
library(tidyverse)

theme_set(theme_bw())

```

```{r paths and parameters}
myinput = "~/../../Corrona LLC/Biostat Data Files - MS/monthly/2024/2024-05-05/final/ms_dmlabimg_2024-05-05.dta"
myoutput = NULL
```


```{r data import}
df.raw <- read_dta(myinput)
```


# Summary

This is to examine the MS lab and imaging data and help make a decision where some variables might have ambiguous "standard units".

# Variables


## Lymphocytes


Here's a summary of the results recorded for Lymphocytes.

```{r}
# create temp data frame for exploration
df <- df.raw %>% 
  filter(event_type_dec == "Lymphocytes") %>% 
  select(id, result)

# get summary statistics
mosaic::favstats(df$result) %>% 
  knitr::kable()


# How many are over 100
summary_df <- df %>%
  mutate(category = ifelse(result <= 100, "Under or Equal to 100", "Over 100")) %>%
  count(category) %>%
  mutate(percentage = n / sum(n) * 100) 


```

Here, only `r summary_df[1,2]` observations out of `r nrow(df)` is over 100, which accounts for `r round(summary_df[1,3],2)`%.

```{r}
summary_df %>% knitr::kable(digits = 2)

```

```{r}
summary_df <- df %>%
  mutate(category = case_when(
    result <= 10 ~ "0 to 10",
    result > 10 & result <= 65 ~ "10 to 65",
    result > 65 & result <= 100 ~ "65 to 100",
    result > 100 ~ "Over 100"
  )) %>%
  count(category) %>%
  mutate(percentage = n / sum(n) * 100)

summary_df %>% knitr::kable(digits = 2)
```

Setting a lower bound of range check to be 10% means we will see `r summary_df[1,2]` observations in the report, which accounts for `r round(summary_df[1,3],2)`% of the observations.


# Session Information
```{r session information, include=T}
sessionInfo()
```

